---
title: "Formulas and Definitions"
description: "Equations and R code"
---

# Parameters and Statistics

| Measure | Population $(\theta)$ | Sample $(\hat{\theta})$ |
|-------------------|--------------------------|---------------------------|
| Total | $\tau = \sum_{i=1}^N y_{i}$ | $\hat{\tau} = \sum_{i=1}^n y_{i}$ |
| Mean | $\mu = \frac{1}{N}\sum_{i=1}^N y_{i}$ | $\bar{y} = \frac{1}{n}\sum_{i=1}^n y_{i}$ |
| Variance | $\sigma^2 = \frac{1}{N}\sum_{i=1}^{N}(y_i-\mu)^2$ | $s^2 = \frac{1}{n-1}\sum_{i=1}^{n}(y_i-\bar{y})^2$ |
| Proportion$^*$ | $p = \mu$ | $\hat{p} = \bar{y}$ |

-   $N$: total population size
-   $n$: total sample size
-   $y_{i}$: value of measurement $y$ on unit $i$
-   $^*$For proportions $y_{i}$ is a binary indicator of success. $y \in {[0,1]}$. I.e. $I(y_{i}=1)$.

## Expected Value and Variance

$$E(\hat{\theta}) = \sum \hat{\theta}*p(\hat{\theta})$$ $$
V(\hat{\theta}) = \sum [\hat{\theta} - E(\hat{\theta})]^2*p(\hat{\theta}) \qquad \mbox{ -or- }\qquad
V(\hat{\theta}) = E(\hat{\theta}^2) - E(\hat{\theta})^2
$$

-   Sums are over all possible values of $\hat{\theta}$.
-   $p(\hat{\theta})$ is the probability of $\hat{\theta}$ occurring.

## Definition: Bias, Variance, Accuracy

$$\mbox{Bias}(\hat{\theta}) = E(\hat{\theta}) - \theta$$

$$V(\hat{\theta})  = E \Big[(\hat{\theta} - E[\hat{\theta}])^2\Big]$$

$$\mbox{MSE}(\hat{\theta}) = V(\hat{\theta}) + [Bias(\hat{\theta})]^2$$

# Sample Weights

The **sampling weights** $w_{i}$are the reciprocal of the inclusion probability $\pi_{i}$

-   **SRS**: $w_{i} =\frac{1}{n}$
-   **SRSWOR**: $w_{i} =\frac{N}{n}$
-   **Stratified**: $w_{hj}=\frac{N_h}{n_h}$
-   **One stage cluster**: $w_{ij}=\frac{N}{n}$

# Simple Random Sample

| Population Quantity | Estimator $(\hat{\theta})$ | Estimated variance of $(\hat{\theta})$ |
|---------------------|----------------------|-----------------------------|
| Mean: $\mu = \frac{\tau}{N}$ | $\frac{\hat{t}}{N} = \frac{\sum_{i\in S} w_{i}y_{i}}{\sum_{i\in S}w_{i}} = \bar{y}$ | $\hat{V}(\bar{y}) = (1-\frac{n}{N})\frac{s^{2}}{n}$ |
| Total: $\tau = \sum_{i=1}^{N} y_{i}$ | $\hat{\tau} = \frac{1}{n}\sum_{i\in S}  w_{i}y_{i} = N\bar{y}$ | $\hat{V}(\hat{\tau}) = N^{2}\hat{V}(\bar{y})$ |
| Proportion: $p$ | $\hat{p} = \bar{y}$ | $\hat{V}(\hat{p}) = (1-\frac{n}{N})\frac{\hat{p}(1-\hat{p})}{n-1}$ |

-   $i \in S$ : Unit $i$ is an element in the sample $S$
-   The **standard error** of the estimate is the square root of the estimated variance.

# Stratified Random Sample

See [section-04](https://sampling-458.netlify.app/notes/cn04-stratified_sampling#theory-of-stratified-sampling-lohr-3.2) for notation. 

| Population Quantity | Estimator $(\hat{\theta})$ | Estimated variance of $(\hat{\theta})$ |
|---------------------|----------------------|-----------------------------|
| Within strata total: $\tau_{h} = \sum_{j} y_{hj}$ | $\hat{\tau}_{h} = N_{h}\bar{y_h}$ | $\hat{V}(\hat{\tau}_{h}) = (1-\frac{n_h}{N_h})N_{h}^2 \frac{s^2_h}{n_h}$ |
| Overall total: $\tau = \sum_{h} \tau_{h}$ | $\hat{\tau}_{str} = \sum_{h}\hat{\tau}_{h}$  | $\hat{V}(\hat{\tau}_{str}) = \sum_{h} \hat{V}(\hat{\tau}_{h})$ |
| Within strata mean: $\mu_{h} = \frac{\tau_{h}}{N_{h}}$ | $\bar{y}_{h} = \frac{1}{n_{h}}\sum_{j} y_{hj}$ |  $\hat{V}(\bar{y}_{h}) = \frac{1}{N^2}V(\hat{\tau}_{h})$ |
| Overall mean:$\mu = \frac{\tau}{N}$  | $\bar{y}_{str} = \sum_{h} \frac{N_{h}}{N}\bar{y}_{h}$ | $\hat{V}(\bar{y}_{str}) = \frac{1}{N^2}\hat{V}(\hat{\tau}_{str})$ |
| Within strata proportion: $p_{h} = \mu_{h}$ | $\hat{p}_{h} = \bar{y}_{h}$  |$\hat{V}(\hat{p}_{h}) = (\frac{n_h}{n_h-1})\hat{p}_h(1-\hat{p}_h)$   |
| Overall proportion: $p = \mu$  |$\hat{p}_{str} = \sum_{h}\frac{N_{h}}{N}\hat{p}_{h}$  | $\hat{V}(\hat{p}_{str}) = \sum_{h} (1-\frac{n_h}{N_h})\big(\frac{N_{h}}{N}\big)^2\Big(\frac{\hat{p}_h(1-\hat{p}_h)}{n_h-1}\Big)$  |

* $\sum_{h}$ is a simplified version of $\sum_{h=1}^{H}$ and $\sum_{j}$ is a simplified version of $\sum_{j=1}^{N_{h}}$

# Cluster Random Sample

⚠️ Note notation change!

* $y_{ij}$: measurement for the $j$th element in $i$th psu
* $N$: the number of clusters (psus) in the population
    * $n$: the number of psus from the sample
* $M_{i}$: number of ssus in psu $i$ in the population
    * $m_i$: the number of ssu's in psu $i$ from the sample
* $M_{0}$: total number of ssus in the population  

| Population Quantity | Estimator $(\hat{\theta})$ | Estimated variance of $(\hat{\theta})$ |
|---------------------|----------------------|-----------------------------|
| Total in psu $i$: $t_{i}=\sum_{j} y_{ij}$ | $\hat{t}_{i} = \frac{M_{i}}{m_i}\sum_{j} y_{ij}$ | - |
| Variance of psu totals : $S_t^2 = \frac{1}{N-1} \sum_{i} \left( t_i - \frac{t}{N} \right)^2$ | $s_t^2 = \frac{1}{n-1} \sum_{i} \left( t_i - \frac{\hat{t}}{N} \right)^2$ | _AKA variance between psu_|
| Overall Total: $t = \sum_{i} t_i$ | $\hat{t} = \frac{N}{n}\sum_{i} \hat{t}_i$ | $N^2 \left(1 - \frac{n}{N}\right) \frac{s_t^2}{n}$  |
| mean in psu $i$: $\mu_{i} = \frac{1}{M_i} \sum_{j} y_{ij}$  | $\bar{y}_{i} = \frac{1}{m_i} \sum_{j} y_{ij}$  | $s_{i}^{2} = \frac{1}{m_i-1} \sum_{j} (y_{ij}-\bar{y}_{i})^2$ (_AKA variance within psu_)|
| Overall mean: $\mu = \frac{1}{M_0} \sum_{i} \sum_{j} y_{ij}$  | $\bar{y} = \frac{\hat{t}}{NM}$  | $\left( 1 - \frac{n}{N} \right) \frac{s_t^2}{nM^2}$ |


------------------------------------------------------------------------

# R commands

This is a quick reference list. See the R companion for the textbook, the package help files,  vignettes or other tutorials listed at the bottom of this page for more information.

::: callout-important
### A note on missing data

If the result of any of the below functions is `NA`, this may indicate that your variable has missing values. Add the `na.rm=TRUE` argument to the `svymean` or `svytotal` functions and that will calculate a complete-case mean/total.
:::


## Analysis

The `survey` package supports the analysis of data collected using complex survey designs. 

### Specify survey design `svydesign`

-   Function call: `svydesign(id = , weights=,  fpc= , data = )`
-   `id` = variable that identifies clusters
-   `weights` = variable that holds the sampling weights
-   `fpc` = finite population correction. Typically defined in the function call.

The argument details can be found on the specified pages in the R companion for the book, and in the respective sections of these notes. 

- SRS: pg 21
- Stratified Random Sample: pg 34
- Cluster sampling: p57

### Estimators

-   **mean**: `svymean(~x, design)`
-   **total**: `svytotal(~x, design)`
-   **proportion**: Use `svytotal` and divide by `N`
-   **CI for the mean or total**: Use `confint()` after calculating the point estimate
-   **CI for proportion**: `svyciprop(~x, design)` Will also print out $\hat{p}$

### Calculating stratum means and variances

* The first argument of `svyby` is the formula for the variable(s) for which
statistics are desired
* `(by=)` is the variable that defines the groups.
* Then list the `design` object
* and the name of the function that calculates the statistics. 
* Set `keep.var=TRUE` to display the standard errors for the statistics.

```{r}
#| eval: false
svyby(~acres92, by=~region, agstrat.dsgn, svytotal, keep.var = TRUE)
```

------------------------------------------------------------------------

# Sampling

The `sampling` package allows you to take random samples from a sampling frame using different sampling frameworks in a reproducible manner. 


1. Setup your sampling frame in a spreadsheet. [This example uses google sheets](https://docs.google.com/spreadsheets/d/13t_2a1nymS-RfAdDN1lq_WrpLD2xjy2rNsol95rD9VA) and the [googlesheets4](https://googlesheets4.tidyverse.org/articles/googlesheets4.html) package. 

2. Import your sampling frame into R. 
```{r}
library(googlesheets4)
frame <- read_sheet("https://docs.google.com/spreadsheets/d/17bg__F6Cq0zBnbPtMBsNCKNM-pyybVnhujvI2J66n_4")
```

3. Use functions from the `sampling` package to draw random samples according to your design. See the links for more details on what the arguments mean. 

* [SRSWOR](https://sampling-458.netlify.app/notes/cn03-srs#using-the-sampling-package).

```{r}
library(sampling)
set.seed(12345)
srs.idx <- srswor(4, length(frame$unit_id)) 
getdata(frame, srs.idx)
```

* [Stratified](https://sampling-458.netlify.app/notes/cn04-stratified_sampling#using-r)
```{r}
library(dplyr)
frame <- frame %>% arrange(group) # sort first
strata.idx <- sampling::strata(data = frame,      # data set
                 stratanames = "group", # variable name
                 size = c(2,3,2,1,2),      # stratum sample sizes     
                 method = "srswor")     # method for selecting within strata
getdata(frame, strata.idx)
```

* [Cluster]()

**One stage cluster**
```{r}
onestage.idx <- sampling::cluster(data=frame,         # Data set
                  clustername = "group",  # variable name containing clusters 
                  size = 3,               # number of clusters
                  method = "srswor",      # how to draw clusters 
                  description = TRUE)     # show descriptive output
getdata(frame, onestage.idx)
```


**Two stage cluster**
```{r}
mstage.idx <- sampling::mstage(data=frame, 
                 stage = c("cluster", ""),  # sampling method for each stage, blank means SRS
                 varnames = list("group", "unit_id"),  # variable names for each stage
                 size = list(3, c(5,5,5)), # 3 psus, 5 ssus from each psu
                 method = c("srswor", "srswor"))

getdata(frame, mstage.idx)
```


------------------------------------------------------------------------

# Vignettes and handbooks

-   [`survey` package vignette](http://r-survey.r-forge.r-project.org/survey/)
-   [The Epidemiologist R handbook](https://epirhandbook.com/en/survey-analysis.html)
-   [Blog post by Zachary Hertz from 2021](https://zacharylhertz.github.io/posts/2021/06/survey-package)
