---
title: "Project"
format: 
  html: default
---

# Project Overview: Shaping the Future of Math & Data Science

This semester, you will go on a real-world project that will directly impact the future of the Mathematics and Data Science programs at Chico State. Working collaboratively as a class, you will design, implement, and analyze an alumni survey to gather valuable insights about their experiences and career paths. This is your chance to apply the sampling and statistical analysis methods you've learned in a practical and meaningful way.

This project isn't just about crunching numbers; it's about teamwork, communication, and project management. As a class you will be responsible for:

-   Defining Roles & Responsibilities: Assign yourselves specific roles within the project (e.g., survey design team, data analysis team, communication lead).
-   Task Management & Timelines: Establish clear timelines and track progress to ensure successful completion of each project phase.
-   Stakeholder Communication: Get stakeholder (Math Department faculty, Career center) input in the questions being asked.
-   Design & Implementing Multiple Sampling Frameworks: Implement multiple sampling methods (e.g., simple random, stratified, cluster) to select representative samples of alumni.
-   Data Collection & Analysis: Collect survey responses, clean the data, and perform appropriate statistical analyses to answer key research questions. You will also need to compare estimates obtained across the different sampling frameworks to assess their impact on the results.
-   Final Presentation: Present your findings and recommendations and insights to the Math Department or the assessment committee.

This is a full class collaborative effort, requiring everyone's active participation and contribution. Every student should take lead on at least one aspect or activity. Your collective efforts will provide valuable data to inform curriculum development, program improvements, and future initiatives within the Mathematics & Statistics departments. Embrace this opportunity to make a real difference!

# Milestones

1.  **Define Project Goals and Stakeholder Input:** Documented agreement on the survey's objectives and how the results will be used, based on input from stakeholders.
2.  **Literature Review and Best Practices Research:** Review of existing alumni surveys and research on best practices in survey design and implementation.
3.  **Create Sampling Frame:** A list of the target population from which the sample will be drawn.
4.  **Develop & Test Questionnaire:** Draft questionnaire, pilot testing, and revisions based on feedback.
5.  **Design Sampling Strategy:** Detailed plan for selecting the sample, including justification for the chosen method and sample size calculations.
6.  **Collect Data:** Completed survey responses. Include metrics on response rate and efforts to address non-response bias.
7.  **Data Cleaning:** Prepare the data for analysis. Screen for outliers, missing or invalid responses. Create necessary metrics. Create basic tables and summaries for each measures.
8.  **Data Analysis:** Create analysis plan, conduct analysis and compile results, create visualizations.
9.  **Final Presentation:** Presentation to stakeholders
10. **Final Report:** Comprehensive written report documenting all aspects of the project.

# Parameters to be Estimated

* Average salary expectations for 1, 2, and 5 years.
* Average time taken by alumni to secure their first job
* Proportion of alumni who found their degree valuable
* Proportion of alumni working in different career sectors
* Proportion of alumni who were accepted to or attended graduate school (mathematics oriented) or with a terminal degree
* Proportion of alumni who felt they were prepared


# Sample Size Determinations

Since most of the parameters to estimate are proportions we will determine the sample size under that framework. Our sampling frame (population size) is $N=300$. 

* We have no prior knowledge of the estimated proportions, so $\hat{p} = \hat{q} = .5$. We will use this for our estimate of $S$. 
* We will set our confidence to be 95%, thus $\alpha = .05$
* We will allow our margin of error to vary between 1% and 25%

I created a graph to plot the sample size under an SRS framework using equation 2.30, 2.31 as a function of the margin of error `e`. We can use this to make a determination of a general total sample size that we will aim for in each of the sampling designs. 

```{r}
srs.n <- function(e){
  n0 <- 1.96^2 / (4*e^2)
  n <- n0 / (1+n0/300)
  return(n)
}
e <- seq(.01, .25, by = .01)
plot(e, srs.n(e), type = "l")

srs.n(seq(.05, .25, by = .05))

```


# Sampling Design

Multiple sampling designs were developed as described below. 

Read in Sampling frame
```{r}
library(sampling)
library(tidyverse)
```

```{r, echo = -1}
frame.url <- "https://docs.google.com/spreadsheets/d/1Mvs2EjZ_eqki9Nv9PlOyl1iGEEEAq5knXh4cz_y3FIc"
sampling.frame <- googlesheets4::read_sheet(frame.url)
```

The population size is determined by the size of our sampling frame, and the sample size was determined using the graph above and a class discussion. 

```{r}
N <- NROW(sampling.frame)
n <- 150
```

## SRS


## Stratified on Year





## Stratified on Major

We hypothesize that students who studied different focuses will be pursuing different types of careers and making different kinds of salaries. Students who double or triple majored will also not have the same experiences as those who share one of their majors. We expect that by stratifying by major, with those with more than one major in a separate stratum, we should be able to expect widely different stratum means for variables like annual income. Since different majors have different course requirements, their feedback on the university will also be different by major. Since we are interested in multiple variables and the observation cost in each stratum will be the same, in this case zero, proportional allocation makes sense as a choice.


First we create a new variable with either a student’s major or if they had multiple majors stored. **Will Stratify by this variable**
```{r}
sampling.frame$stratamajor <- sampling.frame$`Major 1`
sampling.frame$stratamajor[!is.na(sampling.frame$`Major 2`)] = "Multiple Majors"

sampling.frame.sort.major <- sampling.frame %>% arrange(stratamajor)
```

Let `p` be the proportion of total population sampled. We will calculate size of samples drawn from each strata as .... _need equations/info here_


```{r}
p <- n/N

strata.size <- table(sampling.frame.sort.major$stratamajor) |> data.frame()
strata.size$n_h <- round(strata.size$Freq * p)
```

```{r}
set.seed(12345)
strata.major.idx <- sampling::strata(data = sampling.frame.sort.major, 
                              stratanames = "stratamajor", 
                              size = strata.size$n_h, 
                              method = "srswor")
strata.major.chosen <- getdata(sampling.frame.sort.major, strata.major.idx)
```

## Cluster: Same year same major

There is no ideal way to cluster with the information we have in our sampling frame. Clustering by any one demographic variable such as major, or year makes it likely that we will miss a section of the population that will differ from the ones selected significantly. To minimize that, we will have smaller clusters by grouping together students of the same major and the same graduation year. Students with multiple majors will be treated as all having the same “major” like in our ST1 plan above. 

### One Stage Cluster:

Basic organizing
```{r}
# renaming to cluster for simplicity
sampling.frame$cluster <- sampling.frame$"Expected \nGrad Year" 

# getting unique cluster ids based on year
(all_clusters <- unique(sampling.frame$cluster))

# viewing frequencies of years
(table <- table(sampling.frame$cluster) )

# picking my sample size, will round to the nearest whole number
weighted.mean(table)
```

NOTE: the number of sampled clusters is based on the average of 30 grads per year to get roughly 150 individuals in the final total sample.

```{r}
M <- length(unique(sampling.frame$cluster)) # total number of clusters (PSUs, aka GROUPS) in population
m  <- 5 # the number of sampled clusters
```

Sampling from clusters:
```{r}
set.seed(2025)

# sampling the clusters
(sampled_clusters <- sample(all_clusters, size = m, replace = FALSE))

# new dataframe of just sampled clusters
cluster_sample <- sampling.frame[sampling.frame$cluster %in% sampled_clusters, ]

```

> Need to confirm/add weights using `sampling` package. 

### Two Stage Cluster: 


## Create sampling indicators and write back to Google sheets. 

Each method will randomly sample a different group of records There undoubtedly will be substantial overlap. To keep track of which records were chosen in each design, we create binary indicators for each design and add those back to the sampling frame in Google Sheets.  

```{r, eval=FALSE}
sampling.frame$SRS <- 0
sampling.frame$Strata_major <- ifelse(sampling.frame$FIXED_ID %in%
                                        strata.major.chosen$FIXED_ID, 1, 0)
sampling.frame$Strata_year <- 0
sampling.frame$Cluster1 <- ifelse(sampling.frame$FIXED_ID %in%
                                        cluster_sample$FIXED_ID, 1, 0)
sampling.frame$Cluster2 <- 0

sample.these <- sampling.frame %>% 
  select(SRS:Cluster2)
head(sample.these)

googlesheets4::range_write(frame.url, data = sample.these, 
                           range = "P1:T299", 
                           col_names = TRUE)
```



