---
title: "Statistical Estimation"
description: "We review some of the introductory concepts of statistics that will be a pre-requisite for the remainder of the course."
date: 2/3/23
---

# Introduction

In this section we review the basic concepts underlying the selection of an estimator of a population parameter, the method for evaluating its goodness, and the concepts involved in interval estimation. Because the bias and the variance of estimators determine their goodness, we need to review the basic ideas concerned with the expectation and variance of a random variable.

:::{.callout-note appearance=minimal}
_Reminder: Most people come into this class with different pre-requisite classes and instructors. This recap helps ensure that we are all on the same page with concepts and notation that may be new for some._
:::

# Populations and Samples

Statistical inference centers around using information from a sample to understand what might be true about the entire population of interest. If all we see are the data in the sample, what conclusions can we draw about the population? How sure are we about the accuracy of those conclusions? 

On April 29, 2011, Prince William married Kate Middleton in London. The Pew Research Center reports that 34% of US adults watched some or all of the royal wedding. How do we know that 34% of all US adults watched? Did anyone ask **you** if you watched it? In order to know for sure what proportion of US adults watched the wedding, we would need to ask **all** US adults whether or not they watched. This would be very difficult to do. As we will see, however, we can estimate the _population_ proportion _parameter_ quite accurately with a _sample statistic_, as long as we use a random sample. In the case of the royal wedding, the estimate is based on a poll using a random sample of 1006 US adults.


* _Statistical inference_: The process of drawing conclusions about the entire population based on the information in the sample.
* _Parameter_: A number that describes the entire population. $\mu$, $p$, $\tau$
* _Statistic_: A number calculated from a sample. $\bar{x}$, $\hat{p}$, $\hat{\tau}$. 

Generally our goal is to know the value of the population parameter exactly but this usually isn't possible since we usually cannot collect information from the entire population. 

Instead we can select a sample from the population, calculate the quantity of interest for the sample, and use this sample statistic to estimate the value for the whole population.


:::{.callout-warning}
### :star: You try it
The US Census states that 27.5% of US adults who are at least 25 years old have a college bachelor's degree or higher. Suppose that in a random sample  of $n$=200 US residents who are 25 or older, 58 of them have a college bachelor's degree or higher. 

What is the population parameter? What is the sample statistic? Use correct notation for your answer.
:::


The value of a statistic for a particular sample gives a _point estimate_ of the population parameter. If we only have the one sample and don't know the value of the population parameter, 

> this point estimate is our best estimate of the true value of the population parameter.


**Variability of Sample Statistics**: We usually think of a parameter as a fixed value while the sample statistic is considered a random variable. 

:::{.callout-note appearance=minimal}
### :question: :busts_in_silhouette: Turn and talk
Why is a sample statistic such as a sample mean $\bar{x}$ considered a random variable?
:::


Along with the point estimate we also want to know how accurate we can expect the point estimate to be. In other words, if we took another random sample of the same size from the population, is the point estimate from this new sample likely to be similar to the first point estimate or are they likely to be far apart.

:::{.callout-tip icon=true}
### Example: Enrollment in Graduate Programs in Statistics
Graduate programs in statistics often pay their graduate students, which means that many graduate students in statistics are able to attend graduate school tuition free with an assistantship or fellowship. There are 82 US statistics doctoral programs for which enrollment data were available. The data set `StatisticsPhD` lists all these schools together with the total enrollment of full-time graduate students in each program in 2009.

:question: What is the average full-time graduate student enrollment in US statistics doctoral programs in 2009? 
:::   


[I don't feel like loading packages in this page so I'm using the `::` shortcut method to access the `here` and `read_csv` functions to load in the data.]{.aside}

```{r}
stat.phd <- readr::read_csv(here::here("data", "StatisticsPhD.csv"))
head(stat.phd) #always look at your imported data to check for import errors
mean(stat.phd$FTGradEnrollment)
```

Based on the data set, the mean enrollment in 2009 is 53.54 full-time graduate students. Because this is the mean for the entire population of all US statistics doctoral programs for which data were available that year, we have that $\mu=53.54$ students.

:star: Take a random sample of 10 programs from the data file and calculate the mean.

```{r}
my.sample.programs <- sample(stat.phd$FTGradEnrollment, size=10)
mean(my.sample.programs)
```

:::{.callout-note appearance=minimal}
### :question: :busts_in_silhouette: Turn to your neighbor and discuss the following questions

* Did everyone get the same sample mean?
* Does your sample mean exactly equal the population mean?
* If you took another sample of 10, would you get the same sample mean? Why?
* If we created a histogram of all our sample means, what would it look like? Where would it be centered at? What is the spread of the histogram?
:::

Knowing the behavior of of repeated sample statistics (like the mean in the prior example) is critically important. Let's dig into this a little more by repeating this sampling experiment many times. 

[:cry: Getting an error? Take a step back and first use a n=5, and DONT assign it to an object. Visually see with your eyes what the result of the `replicate` function gives you. Then once that is working, you can increase the number of replications, assign it as an object and explore the object.]{.aside}

```{r}
many.means <- replicate(n=100, {
  my.sample.programs <- sample(stat.phd$FTGradEnrollment, size=10)
  mean(my.sample.programs)
})
```

Let's visualize the behavior of the many sample means. 

```{r}
hist(many.means)
summary(many.means)
```

Characteristics of this distribution:  

* _Shape_: The distribution of average enrollment isn't quite normal, there seems to perhaps be two peaks? 
* _Center_: The average enrollment is `r round(mean(many.means),2)`
* _Spread_: Average enrollment ranges from `r range(many.means)`. 


:tada: We have just examined a _sampling distribution_ of the sample mean. 

# Sampling Distribution

A _sampling distribution_ is the distribution of sample statistics computed for different samples of the same size from the same population. A sampling distribution shows us how the sample statistic varies from sample to sample.

:::{.callout-note appearance=minimal}
### :question: :busts_in_silhouette: Turn to your neighbor and discuss the following questions

* What information might be important to get from a sampling distribution?
* What would happen if we increased our sample size (_not the number of replicates_) to $n=20$ rather than $n=10$. What about the sampling distribution would change?
* What would happen if we increased the number of replications (holding $n$ constant) to 1000 instead of 100? What about the sampling distribution would change?
:::

# Confidence Intervals

In general, it is usually not enough to just give a point estimate when estimating a population parameter. Why?


## Standard Error

The _standard error_ of a statistic is the standard deviation of the sample statistic. It tells us how much the sample statistic will vary from sample to sample. In situations like above where we can examine simulated values of the sample statistic we can estimate the standard error by taking the sample standard deviation of the simulated sample statistics. In other situations we can use mathematical formulas to calculate the standard error.

:::{.callout-tip icon=true}
### Example Grad program example cont. 
Estimate the standard error for the mean enrollment in statistics PhD programs for a sample size of 10 and also a sample size of 20. 
:::

[I used the base pipe `|>` here to pass the results of the replicate function into the `sd()` function]{.aside}

```{r}
sd(many.means) #because the example above already had n=10

replicate(n=100, {
  my.sample.programs <- sample(stat.phd$FTGradEnrollment, size=20)
  mean(my.sample.programs)
}) |> sd()
```


When the distributions are relatively symmetric and bell-shaped, the 95% rule tells us that approximately 95% of the data values fall within two standard deviations of the mean. Applying the 95% rule to sampling distributions, we see that about 95% of the sample statistics will fall within two standard errors of the mean. We use this rule many times to form what we call a 95% confidence interval which gives us a range for which which we are quite confident that captures the true parameter we are trying to estimate.

When using a formula to calculate an approximate 95% confidence interval, use $2*SE$ as the margin of error. 

:::{.callout-tip icon=true}
### CI for PhD program enrollment
Based on our example, what would be a 95% confidence interval for $\mu$ the true mean total enrollment for PhD programs in statistics.
Interpret this confidence interval in context of the problem.
:::

[Wrapping the entire line of code in `()` will execute that code AND print out the results. This lets me see the results in the rendered document AND store the results in an object to call later, like in my sentence response below. ]{.aside}

```{r}
(LCL <- mean(many.means) - 2*sd(many.means))
(UCL <- mean(many.means) + 2*sd(many.means))
```

[Every time I compile these notes, I draw a different sample and will get a different numbers. To avoid conflicts in my written response, and what the code shows, I use _inline R code_ here. See the RStudio _Help --> Markdown Quick Reference_ for more information.]{.aside}

> We can be 95% confident that the true mean total enrollment for PhD programs in statistics is covered by the interval (`r paste0(round(LCL,1), " , ", round(UCL,1))`). 



# Summarizing Information in Populations and Samples

## The Infinite Population Case

The first thing that we do with data is to summarize it with graphs or numerical summaries. 
[_Empirical_: Relying on or derived from observation or experiment.]{.aside}

* Histograms or relative frequency tables can be used to try to make an empirical estimation of the shape of the distribution. 
* Once a relative frequency distribution has been established for a population, we can use probability arguments to calculate numerical summary measures such as the mean, variance, and standard deviation. 
* The numerical measures used to summarize the characteristics of a population are defined as **expected values** of $y$. 

:::{.callout-important}
### Definition: Expected Value
$$
E(y)=\sum_{y}y*p(y)
$$
where the summation is over all the values of $y$ for which $p(y)>0$.  
$p(y)$ is the probability of each value of $y$.
:::


:::{.callout-tip icon=false}
### Example calculation of $E(y)$ from a discrete probability distribution
| y   | p(y) |
|-----|------|
| 1   | .1   |
| 2   | .5   |
| 3   | .4   |

$E(y)=\sum_{y}yp(y) = (1)(.1) + (2)(.5) + (3)(.4) = 2.3$
:::

:::{.callout-important}
### Definition: Population Mean
$$E(y)=\mu$$ 
where $y$ is a value of a single measurement chosen at random from the population.
:::


:::{.callout-important}
### Definition: Population Variance & Standard Deviation

$$\sigma^{2} = V(y)=E(y-\mu)^{2}=\sum_{y}\left(y-\mu\right)^{2}*p(y)$$

The variability of measurements in a population can be measured by the **variance** which is defined as the average squared deviation from the mean between a randomly selected measurement $y$ and its mean value:

The standard deviation is defined as the square root of the variance, and is denoted $\sigma$.
:::


We can also calculate these measurements directly from the sample measurements. 

In statistical studies, the population of interest consists of _unknown_ measurements; hence, we can only speculate on the relative frequencies ($p(y)$). In order to gain information about a population we take a sample of $n$, $y_{1},y_{2},...,y_{n}$. We can then infer characteristics about the population. Given our sample data, $y_{1},y_{2},...,y_{n}$, we can calculate the mean and variance:

$$\bar{y} = \frac{1}{n}\sum_{i=1}^{n}y_{i}$$
$$s^{2} = \frac{1}{n-1}\sum^{n}_{i=1}\left(y_{i}-\bar{y}\right)^{2}$$

:::{.callout-warning icon=false}
### :star: You try it
When a few data points are repeated in a data set, the results are often arrayed in a frequency table. For example, a quiz given to each of 25 students was graded on a four-point scale (0,1,2,3), 3 being a perfect score. 

Results: 16 students scored a 3, 4 students scored a 2, 2 students scored a 1, and 3 scored a 0.

Show how the average score can be calculated using both frequencies and proportions, and calculate the standard deviation of $X$. 
:::


:::{.callout-warning icon=false collapse="true" appearance=minimal}
### Solution
| Score (X) | Frequency | Proportion |
|-----------|-----------|------------|
| 3         | 16        | .64        |
| 2         | 4         | .16        |
| 1         | 2         | .08        |
| 0         | 3         | .12        |

$\bar{x}$ = 2.32, s = 1.07

```{r}
(x <- rep(3:0, c(16, 4, 2, 3)))
mean(x)
sd(x)
```
:::

## The Finite Population Case (Lohr Ch 2.2)

The previous section develops results for random sampling from a population considered to be infinite. In such situations, each sampled element has the same chance of being selected and the selections are independent of one another. 

* Most sampling problems don't live in an infinite world, but the population is usually finite although it may be quite large. 
* In addition, we may want to take into consideration varying the probabilities with which the units are sampled. 

:::{.callout-tip icon=true}
### Example: Estimate total number of job openings. 
Suppose, for example, we want to estimate the total number of job openings in a city by sampling industrial firms from within that city. Many of these firms will be small and a few firms will be large. In a random sample _of firms_, the size of the firm is not taken into account. However, the number of job openings will be directly dependent on the size of the firm. Thus, we might improve the sample if large firms are more likely to be included in the sample. 
:::

* Suppose the population consists of the set of elements $\{u_{1},u_{2},...,u_{N}\}$ and a sample of $n$ elements is to be selected with replacement independently of one another.
* Let $\{\delta_{1},\delta_{2},...,\delta_{N}\}$ represent the respective probabilities of selection for the population elements. 
    - $\delta_{i}$ is the probability that $u_{i}$ is selected on any one draw. 
    - for the case of random sampling with replacement, each $\delta_{i} = 1/N$

[Note: The mathematical definition of unbiased is not the same thing as selection biased discussed earlier. This definition will be discussed below.]{.aside}
An unbiased estimator of the population total, $\tau$, is given by

$$
\hat{\tau}=\frac{1}{n}\sum^{n}_{i=1}\frac{y_{i}}{\delta_{i}}
$$

When $\delta_{i} = 1/N$, we can rewrite this as 

$$
\hat{\tau} = \frac{1}{n}\sum^{n}_{i=1}\frac{y_{i}}{(1/N)} = \frac{N}{n} \sum^{n}_{i=1}y_{i} = N\bar{y}
$$

The above estimator is an unbiased estimator for the population total. It is in the best interest to select the $\delta_{i}$ values in such a way that the variance is as small as possible. The best practical way to choose the $\delta_{i}$ is to choose them proportional to a known measurement that is highly correlated with $y_{i}$. 

:::{.callout-tip icon=true}
### Example: Job openings cont. 
Our universe only consists of these $N=4$ firms: A, B, C and D. 

| Firm | Jobs ($x$) | Size of firm | $\delta_{i}$ |
|------|------------|--------------|--------------|
| A    | 3          | 70           | 70/580       |
| B    | 10         | 90           | 90/580       |
| C    | 25         | 120          | 120/580      |
| D    | 61         | 300          | 300/580      |
:::

The population total of job openings is $\tau= 99$. But of course we don't know this, so we want to take a sample of $n=2$ firms and count the nubmer of jobs at each firm $x$ as a way to estimate this total. 

If all firms have equal probability of being selected, then the subset of firms $\{A, B\}$ have the same chance of being selected as any other subset of two firms such as $\{A, C\}$. However, we may want to select firms with probabilities in proportion to their workforce ($\delta_{i}$).  

Without adjustment the estimate of the total job openings based on the $\{A, B\}$ sample would be: 
$$
\hat{\tau} = N\bar{X} = 4*(13/2) = 26
$$


If we adjust $y_{i}$ based on the probability of selecting firms A and B, :
$$
\hat{\tau} = \frac{1}{2}\Big[\frac{3}{70/580} + \frac{10}{90/580}\Big]  = 44.7
$$

# Goodness of Estimation

## Sampling Distribution


## Bias, Precision, and Accuracy

