---
title: "Simple Random Sampling"
description: "SRS is the basic form of probability sampling, and serves as the basis for more complicated forms."
date: 2/24/25
format: 
  html: 
    toc: true
    toc_float: true
execute:
  warning: false
  message: false
---

::: {.callout-note appearance="minimal"}
These notes use functions from the `sampling`, `survey` and `knitr` packages. See the [formulas](https://sampling-458.netlify.app/notes/formulas.html) page for links to vignettes and handbooks. [The best way to learn a new package is to reference the help file and vignette often for examples.  The R Companion for the book (free pdf from book website) is also very helpful. ]{.aside}
:::

```{r, echo=FALSE}
library(sampling);library(survey)
```

# Introduction

This section will aim to answer the following questions: 

* How can we draw a simple random sample of observations from a data set both with, and without replacement: 
* What is the finite population correction (fpc) and why do we need it? 
* What are sampling weights, why do we need them, and how are they created? 
* How do we calculate parameter estimates from an SRS that account for both the fpc and sampling weights? 

# Drawing a Simple Random Sample (Lohr Ch 2.3)

Recall there are two ways to draw simple random samples, with and without replacement.

::: callout-important
### Definition: Simple random sample (SRSWR) with replacement:

A SRSWR of size $n$ from a population of size $N$ can be thought of as drawing $n$ independent samples of size 1. Each unit has the same probability of selection: $\delta_{i} = \frac{1}{N}$

The procedure is repeated until the sample has $n$ units, which may include duplicates.
:::

::: callout-important
### Definition: Simple random sample (SRS) without replacement:

A SRS of size $n$ is selected so that every possible subset of $n$ distinct units in the population has the same probability of being selected as the sample. There are $\binom{N}{n}$ possible samples, resulting in a selection probability for an individual unit $\delta_{i} = \frac{n}{N}$. (See Lohr 2.10 and Appendix A for derivation)
:::

## Intentionality in sampling

Random does not mean haphazard, contrary it's actually quite intentional. Avoid selecting a sample that you "feel" is random or representative of the population. These practices can lead to bias and lack of generalizability.

To avoid the haphazard nature of "blindly choosing", or worse looking at what was sampled and changing it because "it doesn't look random enough", we use techniques that leverage pseudo-random number generating algorithms.


## Drawing a SRS using a computer

::: callout-important
### Process 

1.  Generate a list of all observational units in the population (sampling frame).
2.  Assign each observational unit a unique number, from 1 to the size of the sampling frame $N$.
3.  Use a computer to draw $n$ numbers from 1 to $N$ without replacement.
4.  Subset the data to keep only the selected rows.
:::

:::{.callout-tip icon=true}
### Example: Sampling Statistics PhD programs. 
To _demonstrate_ this example, lets only grab the first 10 programs in the StatisticsPhD data set. From that we'll draw a sample of $n=4$ programs out of these 10. 

```{r}
stats <- readr::read_csv(here::here("data", "StatisticsPhD.csv"))
(stats10 <- stats[1:10,])
```
:::

### Using the `sample()` function

**without replacement**
```{r}
pop.idx <- 1:NROW(stats10)  # Steps 1, 2
(idx <- sample(pop.idx, 4)) # Step 3 
stats10[idx,]               # Step 4
```

* Rows `r idx` were chosen. 

**with replacement**
```{r}
set.seed(4134) # again, for demonstration of duplicate records
(idx.with.dups <- sample(pop.idx, 4, replace=TRUE)) #3 
stats10[idx.with.dups,] #4
```

* Rows `r idx.with.dups` were chosen. 
* Carnegie (row 4) was sampled twice. 

## Using the `sampling` package

* The functions `srswr(n,N)` and `srswor(n,N)` draw SRS with, and without replacement respectively. 
* Each take two arguments: $n$ the sample size, and $N$ the population size. 
    - The vector returned is a vector of length $N$ that indicates how many times that position in the vector is selected.  
    Then the `getdata()` function is used to extract the values from the population the indicated number of times.

**without replacement**

```{r}
(choose.these.wor <- srswor(4, 10)) # Steps 1,2,3
getdata(stats10, choose.these.wor)  # Steps 4
```


**with replacement**
```{r}
set.seed(4134)
(choose.these.wr <- srswr(4,10)) 
getdata(stats10, choose.these.wr)
```

* Again, notice Columbia was chosen twice when sampling with replacement. 
* Interestingly, the result from sampling without replacement added a column called `ID_unit` on it that denotes the row it was sampled from. 

:::{.callout-warning icon=false}
### You try it 

The U.S. government conducts a Census of Agriculture every five years, collecting data on all farms (defined as any place from which \$1000 or more of agricultural products were produced and sold). The file `agpop.csv` (textbook data) contains historical information from 1982, 1987, and 1992 on the number of farms, total acreage devoted to farms, number of farms with fewer than 9 acres, and number of farms with more than 1000 acres for the population consisting of the $N=3078$ counties and county-equivalents in the United States.

Draw a sample of 300 farms without replacement using both `sample` and `srswor`. Save one of these data frames with the name `ag.srs` for later use. 
:::

:::{.callout-warning icon=false collapse="true" appearance=minimal}
### Solution
Import the data and define our sample and population sizes. 
```{r}
ag <- readr::read_csv(here::here("data", "agpop.csv")) 
N <- NROW(ag)
n <- 300
```

**Using the `sample()` function**

[Get in the habit of opening the data set and visually looking at your process at each step. The best way to learn is to check that you know exactly what was done at each step.]{.aside} Generate list of numbers for each observational unit, then draw 5 numbers without replacement.

```{r}
sampling.frame <- 1:N #1, 2
sample.idx <- sample(sampling.frame, n, replace=FALSE) #3
head(sample.idx)
```

Create a subset data frame with only the rows that were chosen and stored in the vector `get.these`.

```{r}
sample.ag <- ag[sample.idx, ]  #4
head(sample.ag[,1:5]) # only showing first 5 columns as an example
```

**Using the `srswor()` function**

```{r}
set.seed(1067)
srswor.idx <- srswor(n, N)
ag.srs <- getdata(ag, srswor.idx) 
```
:::


# Formulas for Estimation

Below is a table of common statistics and how they are estimated under the SRS framework. This table can also be found on the [formulas](https://sampling-458.netlify.app/notes/formulas.html) page. 

| Measure    | Unbiased Estimate $(\hat{\theta})$         | Estimated variance of $(\hat{\theta})$                             |
|------------------|-------------------------|-----------------------------|
| Mean       | $\bar{y} = \frac{1}{n}\sum_{i\in S} y_{i}$ | $\hat{V}(\bar{y}) = (1-\frac{n}{N})\frac{s^{2}}{n}$                |
| Total      | $\hat{\tau} = N\bar{y}$                    | $\hat{V}(\hat{\tau}) = N^{2}\hat{V}(\bar{y})$                      |
| Proportion | $\hat{p} = \bar{y}$                        | $\hat{V}(\hat{p}) = (1-\frac{n}{N})\frac{\hat{p}(1-\hat{p})}{n-1}$ |

* $i \in S$ : Unit $i$ is an element in the sample $S$

Did you notice something different about the formula for the variances? 

::: callout-important
### Finite Population Correction

$$\Big(1-\frac{n}{N}\Big)$$

The larger % of the population that you include in your sample (sampling fraction = $\frac{n}{N}$), the closer you are to a census, the smaller the variability your estimate will have.
:::

* Most samples that are taken from a very large population, the fpc is close to 1.
* So the variance is more determined by the size of the sample, not the % of the population sampled.

:::{.callout-tip icon=true}
### Example
Calculate the estimated standard deviation of the sample mean $\sqrt{\hat{V}(\bar{y})}$ of the the number of acres devoted to farms in 1992 (variable `acres92`). Interpret this number. 
:::

```{r}
y.bar <- mean(ag.srs$acres92)
s2 <- sum((ag.srs$acres92-y.bar)^2)/(n-1)
(sd.ybar <- sqrt((1-n/N)*(s2/n)))
```

[Inline R code: `prettyNum(sd.ybar, big.mark=',')`]{.aside}
Sample means generated from samples of size `r n` will vary from sample to sample by `r prettyNum(sd.ybar, big.mark=',')` acres. 


:::{.callout-warning icon=false}
### You try it
Using a random sample of 500 farms, estimate the standard deviation of the sample proportion of the number of farms with less than 200,000 acres. 

Hint(s)

* Use the `srswor` function to draw the sample
* create an binary indicator to identify farms meeting the criteria
* use the `mean` function to calculate the sample proportion
* use the formula to calculate the sd of the proportion
:::

:::{.callout-warning icon=false collapse="true" appearance=minimal}
### Solution
```{r}
n2 <- 500
srswor.idx2 <- srswor(n2, NROW(ag))
sample.ag.500 <- getdata(ag, srswor.idx2)

p.hat <- mean(sample.ag.500$acres92<200000)
(s.phat <- sqrt((1-n2/N)*(p.hat * (1-p.hat))/(n2-1)))
```

The estimated proportion of number of farms with less than 200k acres varies by 2.04% from sample to sample. 

:::

:::{.callout-important icon=false collapse="true" appearance=minimal}
### Clean up
As we go through these notes, we create a lot of the same named objects, like `n` and `N`, or `idx` and `idx2`. Every once in a while its a good idea to clean out your global environment to not get confused. Use the `rm()` function to remove everything except the population `ag`, our sample of 300 `ag.srs`, and their respective sample sizes: `n` and `N`. 
```{r}
ls()[!ls() %in% c("ag", "ag.srs", "n", "N")]
```

:::

# Sampling Weights (Lohr Ch 2.4)

Recall that a goal of sampling is to obtain a representative sample, one that is similar to the true unknown population. Thus, conceptually if we duplicate certain units from our sample a certain amount of times, we could "reconstruct" what the population looks like. That is, we could create $w_{i}$ copies of unit $i$ for each unit in the sample. 

::: callout-important
### Definition: Sampling Weight (Design weight)

Inverse of the inclusion/selection probability for unit $i$. 

$$w_{i} = \frac{1}{\delta_i}$$

Also interpreted as the number of population units represented by unit $i$.
:::

In an SRS, each unit has an inclusion probability of $\delta_{i} = \frac{n}{N}$, so the sampling weights are all $w_{i} =\frac{N}{n}$. 

We don't _actually_ make $w_{i}$ copies of record $i$, but use these weights as a multiplier in our estimation calculations. 

|         Population size         |                   Total                    |                   Mean                    |
|:---------------------:|:-----------------:|:-----------------------:|
| $\hat{N} = \sum_{i \in S}w_{i}$ | $\hat{\tau} = \sum_{i \in S}w_{i}y_{i}$ | $\bar{y} = \frac{\hat{\tau}}{\hat{N}}$ |

These weighted estimators are used in all probability sampling designs. 

:::{.callout-tip icon=true}
### Example: Calculating weighted estimates
Estimate the total and average number of acres devoted to farms in 1992 using both weighted and unweighted estimates. Then compare these values to the parameter. 
:::

The sampling weights are $w_{i} = \frac{3078}{300}$ for each unit $i$ in the sample, so we'll add that on as a new column before calculating the weighted estimates. 
[See Lohr Table 2.1 for a nicer visual of the data frame with weights. ]{.aside}
```{r}
ag.srs$wt <- N/n
(N.hat <- sum(ag.srs$wt)) # just to confirm to myself
```

Calculate weighted and unweighted estimates, then the pop parameters. 
```{r}
tau.hat.wt <- sum(ag.srs$acres92*ag.srs$wt)
y.bar.wt   <- tau.hat.wt/N.hat
tau.hat.nowt <- sum(ag.srs$acres92)
y.bar.nowt   <- mean(ag.srs$acres92)
mu  <- mean(ag$acres92)
tau <- sum(ag$acres92)
```

Package it in a data frame for easier viewing. 
```{r}
data.frame(
  Measure = c("Total", "Mean"), 
  Parameter = c(tau,mu),
  Unweighted = c(tau.hat.nowt, y.bar.nowt), 
  Weighted = c(tau.hat.wt, y.bar.wt)
  ) |> knitr::kable(align = 'lccc', 
             caption = "Comparing weighted and unweighted estimates")
```

Note that when we have an SRS with equal sampling weights $w_{i} = \frac{N}{n}$, the weighted mean is equal to the unweighted mean. 

:::{.callout-warning icon=false}
###  You try it
Calculate the proportion of farms with less than 200k acres, with, and without weights. Compare to the population proportion. 
:::

:::{.callout-warning icon=false collapse="true" appearance=minimal}
### Solution
```{r}
ag.srs$lt200k <- 1*(ag.srs$acres92<200000)
(p.hat.nowt <- mean(ag.srs$lt200k))
(p.hat.wt   <- sum(ag.srs$lt200k * ag.srs$wt)/N)
(p          <- mean(ag$acres92<200000))
```
:::

# Analysis using the `survey` package (Lohr Ch 2.6)

Survey designs are specified using the `svydesign` function. The main arguments to the the function are id to specify sampling units, weights to specify sampling weights, and fpc to specify finite population size corrections. These arguments should be given as formulas, referring to columns in a data frame given as the data argument.
[Similar to using `group_by` in dplyr, that adds a "flag" to the data set indicting that all subsequent actions are to be done to each group separately, the `svydesign` function adds information to the data set so subsequent actions know how to incorporate measures like weights and strata]{.aside}
```{r}
library(survey)
ag.srs.dsgn <- svydesign(id = ~1, weights = ~wt, fpc = rep(N,n), data = ag.srs)
```

The `id` argument specifies the clusters. We are not using any clustering, so each unit has it's own id. 

## Point and interval estimates

Survey design adjusted estimates can be obtained using `svymean` and `svytotal`. 
```{r}
svymean(~acres92, ag.srs.dsgn)
svytotal(~acres92, ag.srs.dsgn)
svymean(~lt200k, ag.srs.dsgn)
```

Estimates for multiple variables can be obtained at the same time. 
```{r}
svymean(~acres92+farms92, ag.srs.dsgn)
```

You can pass the results to construct confidence intervals. 
```{r}
svytotal(~acres92+farms92, ag.srs.dsgn) |> confint()
```

We can be 95% confident that the interval (1,646,531, 1,937,985) covers the true number of farms in the united states in 1992. 

:::{.callout-warning icon=false}
###  You try it
Using your sample of 300 farms, estimate the _total_ number of farms in the United States in 1987 and the _average_ farm size in acres for the same year. Report both point and interval estimates. 
:::

:::{.callout-warning icon=false collapse="true" appearance=minimal}
### Solution
```{r}
pt.est.t <- svytotal(~farms87 , ag.srs.dsgn)
ci.t <- pt.est.t |> confint() |> prettyNum(big.mark = ",")
pt.est.m <- svymean(~acres87, ag.srs.dsgn)
ci.m <- pt.est.m |> confint() |> prettyNum(big.mark = ",")
```
Basic info:

* Based on this sample of 300 farms, the total number of farms in the US in 1987 is `r prettyNum(pt.est.t[1], big.mark = ",")`, with 95% CI (`r ci.t[1]`,`r ci.t[2]` )
* and the average acres for those farms is `r prettyNum(pt.est.m[2], big.mark = ",")`, with 95% CI (`r ci.m[3]`,`r ci.m[4]`)

Fancy table
```{r}
data.frame(
  Measure = c("Total number of farms", "Average farm size"), 
  Estimate = prettyNum(c(pt.est.t[1], pt.est.m[1]), big.mark = ','),
  LCL = c(ci.t[1], ci.m[1]), 
  UCL = c(ci.t[2], ci.m[2])
  ) |> knitr::kable(align = 'lccc', row.names = FALSE,
             caption = "Point estimates and confidence intervals for the total number of farms in the US in 1987 and the average farm size in that same year.")
```
:::


