---
title: "Stratified Sampling"
description: "A method to use additional information about the population in the survey design. Lohr Ch 3. "
author: DRAFT
date: 3/20/23
execute: 
  error: true
  warning: false
  message: false
---

# What is Stratified Sampling 


## Overview

* A sampling technique in which we divide the desired population into subgroups called _strata_
* Each individual is necessarily in a single strata; no overlap within strata
* All strata consist of the entire population
* An SRS is taken from each strata, then results can be pooled to make inferences about the entire population

### Why Use A Stratified Sample?

* Protection from bad samples; less likely of drawing a non representative sample through pure chance
* Want data of known subgroups (i.e.,  if we are interested in people’s heights, we know that men tend to be taller than women)
* Convenience / cost effectiveness 
– At what point will this compromise the legitimacy of the study? 
* Lowers variance when done properly; variance within each strata is likely lower than the population variance
* you do not have to sample the same fraction of observations in each stratum. If the variability is higher in a certain strata then you can try to reduce the variability of the estimated total by taking more observations from that stratum. 
:::{.callout-tip icon=true}

### Motivating Example - Estimating the average number of farm acres.  

```{r}
library(here); library(tidyverse)
agpop <- read_csv(here("data", "agpop.csv"))
```

In Example 2.6  we took a SRS to find the average number of farm acres per county. Even though our methods of taking an SRS were correct, some areas of the country were over or underrepresented due to the variability in the size of farms in different regions in the country. An example of this would be that western states in the U.S tend to be larger, and therefore tend to have larger farms. 

To try and fix this, we can use a stratified sample can give balance to the sample on the stratifying variable, which in this case would be region that the farms are located. 
:::

To begin, we can divide our data into four regions (these are the strata): 

```{r}
unique(agpop$region) # the four regions and their names
```

Previously we sampled 10% of the population with an SRS, and so to be able to compare our results with our stratified sample, we should sample 10% of the population from each strata through four separate and independent SRS's. 

The process to do this would be to take the four regions, number off the counties that are in the region, calculate what number of counties is 10% of the total number of counties in the region, and select that many numbers from the list of numbered counties. 

> _Details on how to do this in R is covered in a later section. For now we will use the `agstrat` dataset from the book as the stratified sample._ 

```{r}
agstrat <- read_csv(here("data", "agstrat.csv"))
```


:::{.callout-tip icon=true}
### Exploring the distribution of farm acreage in 1992 by region.

```{r}
#| fig-cap: The thick line for each region is the median of the sample data from that region. The upper and lower horizontal lines of the boxes are the 25th and 75th percentiles. 
#| fig-cap-location: top

boxplot(acres92/10^6 ~ region, data = agstrat,
        xlab = "Region", ylab = "Millions of Acres")

```

The Northeast has the lowest median and the smallest variance. The West has the highest median and the highest mean and variance. The distribution of farm acreage appears to to be positively skewed in each region.

:::

:::{.callout-warning icon=false}
### :star: You try it. 
Explore the distribution of farm acreage in 1987 by region. Create a table of summary statistics and a boxplot. Discuss your findings. 
:::

:::{.callout-warning icon=false collapse="true" appearance=minimal}
### Solution
```{r}
agstrat %>%
  group_by(region) %>%
  summarize(y.bar = mean(farms87),
            s2 = var(farms87),
            n. =n())

ggplot(agstrat, aes(x = region, y = farms87, fill = region)) +
  geom_boxplot() +
  labs(title = "Box Plot of Farm acreage in 1987 by Region",
       x = "Region",
       y = "Acreage")
```

:::

:::{.callout-important}
### MISSING INFO
This group needs to add text information here. Interpret your numbers and plots in context of the purpose of the motivating example. 
:::

# Theory & Weights (Lohr 3.2, 3.3)




# Selecting a Stratified Random Sample


## Using Base R


## Using the `strata` function from the `sampling` package

1. Sort the data by the stratification variable
```{r}
agpop.sorted <- agpop %>% arrange(region)
```

2. Calculate the desired within-strata sample sizes $n_{k}$. For example if we want a 10% SRS within strata, we would find $N_{k}$ and multiply it by 0.1. **Round up to the next digit in case of decimal places**
```{r}
(N.k <- table(agpop.sorted$region) %>% as.vector())
(n.k <- ceiling(0.1*N.k))
```

3. Create the sampling index of records to select using the `strata` function with arguments specifying the strata names, strata size, and sampling method within strata. 
```{r}
strat.idx <- strata(data = agpop.sorted, stratanames = "region", 
                    size = n.k, method = "srswor")
```

4. Extract the data records that correspond to the sampling index. 
```{r}
ag.str <- getdata(agpop.sorted, strat.idx)
head(ag.str)
```

5. Calculate the sampling weights using the inclusion probabilities (these were created as the variable `Prob` when you used the `strata` function)
```{r}
ag.str$wt <- 1/ag.str$Prob
head(ag.str)
```

5b. Check that the sampling weights sum to the stratum population sizes. 
```{r}
ag.str %>% group_by(region) %>% summarize(sum.wts = sum(wt))
N.k
```

:::{.callout-important}
### Warning
If they do not sum to the stratum population sizes, you have made a mistake somewhere in the calculations for the weights or strata sizes. 
:::

# Computing Statistics from a Stratified Random Sample

## Setup the information for the survey design. 

### Specify weights

We created the weights in the dataset in the the last step. 

### Specify fpc

We need to create a vector of fpc's that are equal to the strata population size. In a SRS, we used `fpc = rep(N,n)`. This create a vector that contains the value `N`, repeated `n` times. In this case, we need to match the strata population size to each record from that strata. 

Option 1: Create a vector that repeats the values in the vector `N.k`, each `n.k` times. 
:::{.callout-tip icon=true}
#### Example
```{r}
ex.N.k <- c(10,25,50) # example population sizes
ex.n.k <- c(2,5,10)   # example strata sizes
rep(ex.N.k, ex.n.k)   
```

The pop size for stratum 1 is repeated twice (the size of the sample from that stratum), the pop size for stratum 2 is repeated 5 times (the sample size for stat stratum) etc. 
:::

```{r}
option1.fpc <- rep(N.k, n.k)
```

Option 2: Merge the pop sizes onto the sample data using `region` as a joining key
```{r}
(pop.strata.sizes <- table(agpop.sorted$region) %>% data.frame())
pop.strata.sizes <- rename(pop.strata.sizes, region = Var1, popsize = Freq)
ag.str.opt2 <- ag.str %>% left_join(pop.strata.sizes)
```

### Call `svydesign`
```{r}
agpop.str.dsgn <- svydesign(id = ~1, strata = ~region, weights = ~wt, 
                            fpc = ~popsize, 
                            # fpc = option1.fpc # This would be Option 1
                            data= ag.str.opt2)
```

## Calculate mean, SE and CI
Note the degrees of freedom for a CI under a stratified sample is $n-H$, which can be extracted from the `svydesign` object using the `degf` function. 

```{r}
(str.total <- svymean(~acres92, agpop.str.dsgn))
confint(str.total, level = .95, df = degf(agpop.str.dsgn))
```

Compare these to our prior "hand" calculated values (and approximate 95% CI): 
```{r}
tau.hat.str
tau.hat.str - 2*sqrt(V.tau.hat.str)
tau.hat.str + 2*sqrt(V.tau.hat.str)
```


