---
title: "Formulas and Definitions"
description: "Equations and R code"
date: "2023-01-20"
---

# Parameters and Statistics

| Measure        | Population $(\theta)$                             | Sample $(\hat{\theta})$                            |
|------------------|--------------------------|---------------------------|
| Total          | $\tau = \sum_{i=1}^N y_{i}$                       | $\hat{\tau} = \sum_{i=1}^n y_{i}$                  |
| Mean           | $\mu = \frac{1}{N}\sum_{i=1}^N y_{i}$             | $\bar{y} = \frac{1}{n}\sum_{i=1}^n y_{i}$          |
| Variance       | $\sigma^2 = \frac{1}{N}\sum_{i=1}^{N}(y_i-\mu)^2$ | $s^2 = \frac{1}{n-1}\sum_{i=1}^{n}(y_i-\bar{y})^2$ |
| Proportion$^*$ | $p = \mu$                                         | $\hat{p} = \bar{y}$                                |

-   $N$ = total population size

-   $y_{i}$ = Value of measurement $y$ on unit $i$

-   $^*$For proportions $y_{i}$ is a binary indicator of success. $y \in {[0,1]}$. I.e. $I(y_{i}=1)$.

## Expected Value and Variance of estimates

$$E(\hat{\theta}) = \sum\hat{\theta}*p(\hat{\theta}) \qquad \qquad V(\hat{\theta}) = \sum [\hat{\theta} - E(\hat{\theta})]^2*p(\hat{\theta})$$

-   Sums are over all possible values of $\hat{\theta}$.
-   $p(\hat{\theta})$ is the probability of $\hat{\theta}$ occurring.
    -   This is similar to what we're calling $\delta_{i}$, the probability that population unit $i$ appears in the sample. The book uses the symbol $\pi_{i}$

----

# Weighted Estimators for Probability sampling designs

Survey weights can be used in all probability sampling designs and calculated as the reciprocal of the inclusion probability $\delta_{i}$. Weights $w_{i}$ are applied directly to the data value $y$ before use in any formulas. 


|         Population size         |                  Total                  |                  Mean                  |
|:-----------------------:|:-------------------:|:-------------------------:|
| $\hat{N} = \sum_{i \in S}w_{i}$ | $\hat{\tau} = \sum_{i \in S}w_{i}y_{i}$ | $\bar{y} = \frac{\hat{\tau}}{\hat{N}}$ |


* **SRS**: $w_{i} =\frac{N}{n}$
* **Stratified**: $w_{hj}=\frac{N_h}{n_h}$
* **One stage cluster**: is an SRS of clusters.
* **Two stage cluster**: $w_{ij} = \frac{N}{n}\frac{M_{i}}{m_{i}}$
    - Stage 1: $P$($i$th psu selected) = $\frac{n}{N}$
    - Stage 2: $P$($j$th ssu selected | $i$th psu selected) = $\frac{m_i}{M_i}$

----

# Estimators under different Sampling Designs

## Simple Random Sample

| Measure    | Unbiased Estimate $(\hat{\theta})$         | Estimated variance of $(\hat{\theta})$                             |
|------------------|---------------------|---------------------------------|
| Mean       | $\bar{y} = \frac{1}{n}\sum_{i\in S} y_{i}$ | $\hat{V}(\bar{y}) = (1-\frac{n}{N})\frac{s^{2}}{n}$                |
| Total      | $\hat{\tau} = N\bar{y}$                    | $\hat{V}(\hat{\tau}) = N^{2}\hat{V}(\bar{y})$                      |
| Proportion | $\hat{p} = \bar{y}$                        | $\hat{V}(\hat{p}) = (1-\frac{n}{N})\frac{\hat{p}(1-\hat{p})}{n-1}$ |

-   $i \in S$ : Unit $i$ is an element in the sample $S$
-   $(1-\frac{n}{N})$: Finite population correction (fpc)


## Stratified Random Sample

* CI's should use a $t_{n-H}$ distribution instead of using the normal assumption. 

| Measure                   | Population $(\theta)$                                                  | Sample $(\hat{\theta})$                                     |
|------------------|-----------------------------|-----------------------------|
| Sample size in strata $h$ | $N_{h}$                                                                | $n_{h}$                                                     |
| Overall sample size       | $N = \sum^{H}_{h=1} N_{h}$                                             | $n = \sum^{H}_{h=1} n_{h}$                                  |
| Var(y) in strata $h$    | $\sigma_{h}^2 = \frac{1}{N_{h}-1}\sum_{j=1}^{N_{h}}(y_{hj}-\mu_{h})^2$ | $s_{h}^2 = \frac{1}{n_{h}-1}\sum_{j}(y_{hj}-\bar{y}_{h})^2$ |
| Total in strata $h$       | $\tau_{h} = \sum_{j=1}^{N_{h}} y_{hj}$                                 | $\hat{\tau}_{h} = N_{h}\bar{y_h}$                           |                         |
| Overall total             | $\tau = \sum_{h=1}^{H} \tau_{h}$                                       | $\hat{\tau}_{str} = \sum_{h=1}^{H}N_{h}\bar{y}_{h}$         |
| Mean in strata $h$        | $\mu_{h} = \frac{\tau_{h}}{N_{h}}$                                     | $\bar{y}_{h} = \frac{1}{n_{h}}\sum_{j} y_{hj}$              |
| Overall mean              | $\mu = \frac{\tau}{N}$                                                 | $\bar{y}_{str} = \sum_{h=1}^{H} \frac{N_{h}}{N}\bar{y}_{h}$ |
| Proportion in strata $h$ | $p_{h} = \mu_h$ | $\bar{y}_h = \hat{p}_h$ |
| Variance of $p_{h}$ | $\sigma_h^2=\frac{N_h}{N_h-1} p_h(1-p_h)$ | $s_h^2=\frac{n_h}{n_h-1} \hat{p}_h(1-\hat{p}_h)$ |
| Overall proportion | $p_{str} = \sum_{h=1}^{H} \frac{N_h}{N} p_h$ | $\hat{p}_{str} = \sum_{h=1}^{H} \frac{N_h}{N}\hat{p}_h$  |

Estimated variances; 

* Within strata total $\hat{V}(\hat{\tau}_{h}) = (1-\frac{n_h}{N_h})N^2 \frac{s^2_h}{n_h}$  
* Within strata mean  $\hat{V}(\bar{y}_{h}) = \frac{1}{N^2}V(\hat{\tau}_{h})$
* Overall population total $\hat{V}(\hat{\tau}_{str}) = \sum_{h=1}^H (1-\frac{n_h}{N_h})N^2 \frac{s^2_h}{n_h}$
* Population proportion $\hat{V}(\hat{p}_{str}) = \sum_{h=1}^H (1-\frac{n_h}{N_h}) (\frac{N_h}{N})^2 \frac{\hat{p}_h (1-\hat{p}_h)}{n_h-1}$
* Estimated total number of population units having a specified characteristic: $\hat{t}_{\it{str}} = \sum_{h=1}^H N_h \hat{p}_h$

## Cluster Random Sample


::: panel-tabset
## Pop psu level

| Symbol      | Formula                                                          | Description                                    |
|------------|--------------------------|----------------------------------------|
| $y_{ij}$    |                                                                  | measurement for the $j$th element in $i$th psu |
| $N$         |                                                                  | number of clusters (psus) in the population    |
| $M_{i}$     |                                                                  | number of ssus in psu $i$                      |
| $M_{0}$     | $\sum_{i=1}^{N} M_{i}$                                           | total number of ssus in the population         |
| $t_{i}$     | $\sum_{j=1}^{M_{i}} y_{ij}$                                      | total in psu $i$                               |
| $t$         | $\sum_{i=1}^{N} t_{i}= \sum_{i=1}^{N} \sum_{j=1}^{M_{i}} y_{ij}$ | population total                               |
| $S_{t}^{2}$ | $\frac{1}{N-1} \sum_{i=1}^{N} (t_{i} - \frac{t}{N})^2$           | population variance of the psu totals          |

## Pop ssu level

| Symbol           | Formula                                                                  | Description                        |
|-------------------|-----------------------------------|-------------------|
| $\bar{y}_{\mu}$  | $\frac{1}{M_0} \sum_{i=1}^N \sum_{j=1}^{M_i} y_{ij}$                     | population mean                    |
| $\bar{y}_{i\mu}$ | $\frac{1}{M_i} \sum_{j=1}^{M_i} y_{ij} = \frac{t_i}{M_i}$                | population mean in psu $i$         |
| $S^2$            | $\frac{1}{M_0-1} \sum_{i=1}^N \sum_{j=1}^{M_i} (y_{ij}-\bar{y}_{\mu})^2$ | population variance (per ssu)      |
| $S_{i}^{2}$      | $\frac{1}{M_i-1} \sum_{j=1}^{M_i} (y_{ij}-\bar{y}_{\mu})^2$              | population variance within psu $i$ |

## Sample
| Symbol                   | Formula                                                  | Description                                     |
|------------------|-----------------------------|-----------------------------|
| $y_{ij}$ |                                                                |Measurement for$j$th element in $i$ht psu                                             |
|$N$||Number of clusters (psus) in the population|
|$M_i$||Number of ssus in psu $i$|
|$M_{o}$|$\sum_{i=1}^NM_i$|Total number of psus in the popultion|
|$t_i$|$\sum_{j=1}^{M_i}y_{ij}$|Total in psu $i$|
|$t$|$\sum_{i=1}^Nt_i = \sum_{i=1}^N\sum_{j=i}^{M_i}y_{ij}$|Popultion Total|
|$S_i^2$|$\frac{1}{N-1} \sum_{i=1}^N(t_i-\frac{t}{N})^2$|Popultion variance of the psu totals|

:::

## Ratio Estimation in SRS

Let $y_{i}$ be the measure of interest, and $x_{i}$ be the auxiliary variable. In a population of size $N$ the totals for each measure are

$$
t_{y} = \sum_{i=1}^Ny_{i} \qquad t_{x} = \sum_{i=1}^Nx_{i}
$$

and their ratio is calculated as
$$
B = \frac{t_{y}}{t_{x}} = \frac{\mu_{y}}{\mu_{x}}.
$$

## Regression Estimation in SRS

Regression estimation uses the method of ordinary least squares to calculate estimates for the slope $B_1$ and intercept $B_{0}$ of a 'line of best fit' between the outcome measure of interest $y$ and the auxiliary variable $x$. 

$$
y = B_0 + B_1x
$$

where

$$
\hat{B}_1 = \frac{rs_y}{s_x} \qquad \hat{B}_0 = \bar{y} - \hat{B}_1\bar{x}
$$

where $r$ is the sample correlation coefficient between $x$ and $y$, and $s_x$ and $s_y$ are sample standard deviations of $x$ and $y$ respectively. 


----

# R commands 

This is a quick reference list. See the R companion for the textbook, the `survey` package help file or vignette for more information. 

## Specify survey design `svydesign`

* Function call: `svydesign(id = , weights=,  fpc= , data = )`
* `id` = variable that identifies clusters 
* `weights` = variable that holds the sampling weights
* `fpc` = finite population correction. Typically defined in the function call. 


The argument details can be found on the specified pages in the R companion for the book. 

* SRS: pg 21
* Stratified Random Sample: pg 34
* Cluster sampling: 
* Ratio estimates: pg 41
* Regression estimates: pg 44

## Estimators

* **mean**: `svymean(~x, design)`
* **total**: `svytotal(~x, design)`
* **proportion**: Use `svytotal` and divide by `N`
* **CI for the mean or total**: Use `confint()` after calculating the point estimate
* **CI for proportion**: `svyciprop(~x, design)` Will also print out $\hat{p}$
* **Ratio estimate**: `svyratio(~numerator, ~denominator, design)`
* **Regression estimate**: `svyglm(y~x, design)`

## Calculating stratum means and variances

* The first argument of `svyby` is the formula for the variable(s) for which
statistics are desired
* `(by=)` is the variable that defines the groups.
* Then list the `design` object
* and the name of the function that calculates the statistics. 
* Set `keep.var=TRUE` to display the standard errors for the statistics.

```{r}
#| eval: false
svyby(~acres92, by=~region, agstrat.dsgn, svytotal, keep.var = TRUE)
```

# Manage data collection 

Setup a data collection sheet using random sampling from a sampling frame

1. Setup your sampling frame in a spreadsheet. [See this example using google sheets](https://docs.google.com/spreadsheets/d/1L0voUNhw5CsZ3YPWsuTJCBYm963L0zFIhDKS5HUjmms). 
2. Import your sampling frame into R. More info can be found on the [googlesheets4](https://googlesheets4.tidyverse.org/articles/googlesheets4.html) package vignette. 
```{r}
library(googlesheets4)
sheet.url <- "https://docs.google.com/spreadsheets/d/1L0voUNhw5CsZ3YPWsuTJCBYm963L0zFIhDKS5HUjmms"
frame.srs <- read_sheet(sheet.url, sheet = "srs") 
frame.str <- read_sheet(sheet.url, sheet = "stratified") 
frame.clu <- read_sheet(sheet.url, sheet = "cluster") 
```

3. Use functions from the `sampling` package to draw random samples according to your design. See the links for more details on what the arguments mean. 

* [SRS](https://sampling-458.netlify.app/notes/cn04-srs#drawing-a-srs-using-a-computer).

```{r}
library(sampling)
set.seed(12345)
srs.idx <- srswor(8, length(frame.srs$unit_id)) 
getdata(frame.srs, srs.idx)
```


* [Stratified](https://sampling-458.netlify.app/notes/cn06-stratified_sampling.html#using-r)
```{r}
library(dplyr)
frame.str <- frame.str %>% arrange(group) # sort first
strata.idx <- sampling::strata(data = frame.str,      # data set
                 stratanames = "group", # variable name
                 size = c(6,4,5),      # stratum sample sizes     
                 method = "srswor")     # method for selecting within strata
getdata(frame.str, strata.idx)
```

* One stage cluster
```{r}
onestage.idx <- sampling::cluster(data=frame.clu,         # Data set
                  clustername = "group",  # variable name containing clusters 
                  size = 3,               # number of clusters
                  method = "srswor",      # how to draw clusters 
                  description = TRUE)     # show descriptive output
getdata(frame.clu, onestage.idx)
```



* Two stage cluster
```{r}
mstage.idx <- sampling::mstage(data=frame.clu, 
                 stage = c("cluster", ""),  # sampling method for each stage, blank means SRS
                 varnames = list("group", "unit_id"),  # variable names for each stage
                 size = list(3, c(5,5,5)), # 3 psus, 5 ssus from each psu
                 method = c("srswor", "srswor"))

getdata(frame.clu, mstage.idx)
```

See `?mstage` for more examples of multistage sampling. 

4. Make a new spreadsheet containing the sampled rows, and add a column for $y$ the thing you are going to measure on each unit. 

5. Go collect data! 


## Other R help References

-   [`survey` package vignette](http://r-survey.r-forge.r-project.org/survey/)
-   [The Epidemiologist R handbook](https://epirhandbook.com/en/survey-analysis.html)

