---
title: "Simple Random Sampling (Lohr Ch 2.3)"
description: "SRS is the basic form of probability sampling, and serves as the basis for more complicated forms."
date: 2/13/23
---

> This code file uses the `agpop.csv` data that can be downloaded from the textbook resources

# Introduction

There are two ways to draw simple random samples, with and without replacement. 

## With replacement

* The same unit is "put back" into the pool of eligible units to sample from, and thus has a non-zero probability of being selected again.
* Drawing a subset of $n$ units from the population can be thought of as drawing $n$ independent samples of size 1. We use this feature when drawing random numbers from named distributions using the `size` argument. 
[If you want to reproduce your exact numeric answer later, call `set.seed` immediately before any random draw. See the R companion for the text 2.1 for more info.]{.aside}

This code
```{r}
set.seed(12356)
rnorm(10, 0,1) # draw 10 
```
is equivalent to  
```{r}
set.seed(12356)
replicate(10, { # repeat 10 times
  rnorm(1, 0,1) # draw 1
})
```

## Without replacement

The selected unit is _not_ returned, and thus cannot be chosen a second time. This tends to be the preferred method, sampling the same unit multiple times does not provide more information. 

Every possible subset of $n$ distinct units has the same probability of getting sampled. 


# Drawing an SRS

Random does not mean haphazard, contrary it's actually quite intentional. Avoid selecting a sample that you "feel" is random or representative of the population. These practices can lead to bias and lack of generalizability. 

Instead we use techniques that leverage pseudo-random number generating algorithms such as the `sample` function. 
  
  
:::{.callout-important}
### How to draw a SRS 
1. Generate a list of all observational units in the population (sampling frame). 
2. Assign each observational unit a unique number, from 1 to the size of the sampling frame $N$. 
3. Use a computer to draw $n$ numbers from 1 to $N$ without replacement.   
4. Subset the data to keep only the selected rows. 
:::

:::{.callout-tip icon=true}
### Example: Enrollment in Graduate Programs in Statistics

The [last time we saw this example](https://sampling-458.netlify.app/complete_notes/cn03-statistical_foundations_complete.html#variability-of-estimates), we used the `sample()` function to draw 10 values from the single variable `FTGradEnrollment`. Let's revisit this data and take an SRS of the entire dataset.  
:::

Read in the data and see how many record/rows there are. 
```{r}
stat.phd <- readr::read_csv(here::here("data", "StatisticsPhD.csv")) 
NROW(stat.phd) 
```
[Get in the habit of opening the data set and visually looking at your process at each step. The best way to learn is to check that you know exactly what was done at each step.]{.aside}
Generate list of numbers for each observational unit, then draw 10 numbers without replacement. 
```{r}
sampling.frame <- 1:NROW(stat.phd) #1, 2
(get.these <- sample(sampling.frame, 10, replace=FALSE)) #3
```

Create a subset data frame with only the rows that were chosen and stored in the vector `get.these`. Confirm there are only 10 rows in this subset. 
```{r}
sample.stat.phd <- stat.phd[get.these, ]  #4
NROW(sample.stat.phd)
```

For fun, calculate the average enrollment and compare it to the estimate from the last set of notes
```{r}
mean(sample.stat.phd$FTGradEnrollment)
```








:::{.callout-warning icon=false}
### :star: You try it (Lohr Ex 2.6)

The U.S. government conducts a Census of Agriculture every five years, collecting data on all farms (defined as any place from which \$1000 or more of agricultural products were produced and sold). The file `agpop.csv` (textbook data) contains historical information from 1982, 1987, and 1992 on the number of farms, total acreage devoted to farms, number of farms with fewer than 9 acres, and number of farms with more than 1000 acres for the population consisting of the $N=3078$ counties and county-equivalents in the United States. 

Take a random sample of $n=300$ from this data set, then generate a histogram of the variable `acres92`. Compare your results to your neighbor, and to Figure 2.4 in the text. _Set a seed for this example to ensure you get the same result each time you compile!_
:::

> Write a sentence here about how similar, or different your chart is from the class or the text. 





* Sampling Weights (Lohr Ch 2.4)

# Estimating population parameters

## mean
* Expected value, variance, SE, CV, Bounds on error of estimation
* FPC
* Coefficient of variation

## total: 
* Expected value, variance, SE, CV 

## Proportion
Expected value, variance


# Using the `survey` package. 


# Determining sample size 

# Model based theory vs Randomization based theory (Lohr 2.10, 2.11)

# Closing: When should SRS be used
* comparison to systemic sampling? 


# Analysis under SR






