---
title: "Simple Random Sampling (Lohr Ch 2.3)"
description: "SRS is the basic form of probability sampling, and serves as the basis for more complicated forms."
date: 2/20/23
execute: 
  cache: true
---

::: {.callout-note appearance="minimal"}
These notes use functions from the `sampling` package.
:::

```{r, echo=FALSE}
library(sampling)
```

# Introduction

Recall there are two ways to draw simple random samples, with and without replacement.

::: callout-important
### Definition: Simple random sample (SRSWR) with replacement:

A SRSWR of size $n$ from a population of size $N$ can be thought of as drawing $n$ independent samples of size 1. Each unit has the same probability of selection: $\delta_{i} = \frac{1}{N}$

The procedure is repeated until the sample has $n$ units, which may include duplicates.
:::

::: callout-important
### Definition: Simple random sample (SRS) without replacement:

A SRS of size $n$ is selected so that every possible subset of $n$ distinct units in the population has the same probability of being selected as the sample. There are $\binom{N}{n}$ possible samples, resulting in a selection probability for an individual unit $\delta_{i} = \frac{n}{N}$. (See Lohr 2.10 and Appendix A for derivation)
:::

# Intentionality in sampling

Random does not mean haphazard, contrary it's actually quite intentional. Avoid selecting a sample that you "feel" is random or representative of the population. These practices can lead to bias and lack of generalizability.

::: {.callout-note appearance="minimal"}
### :question: :busts_in_silhouette: Turn and talk

Suppose you want to estimate the number of weed clusters of a certain type in a field.

-   What is the population and what would you use for sampling units?
-   How would you select a simple random sample?
-   If a sampling unit is an area such as a square yard, does the size chosen for a sampling unit affect the accuracy of results?
-   What considerations go into our choice of sampling unit?
:::

::: {.callout-note icon="false" collapse="true" appearance="minimal"}
### Class responses

-   Population: All plots of that type of field
-   Sampling unit: The plots of land
-   Yes. Larger plots of land will have more weed clusters
-   Feasibility, size of land, terrain type
:::

To avoid the haphazard nature of "blindly choosing", or worse looking at what was sampled and changing it because "it doesn't look random enough", we use techniques that leverage pseudo-random number generating algorithms such as the functions discussed next

# Drawing a SRS using a computer

Previously we saw how to use the `sample` function to accomplish this. An alternative is to use the functions `srswor` or `srswr` from the `sampling` package. We'll first look at how these functions differ.

When using `sample()` the vector of data that you want to sample from is provided as the first argument, and what is returned is the values in that vector.

```{r}
my.animals <- c("Riley", "Meka", "Toki", "TJ", "DC", "Dodger")
sample(my.animals, 3)
```

The functions `srswr` and `srswor` take two arguments: $n$ the sample size, and $N$ the population size, and what is returned is a vector of length $N$ that indicates how many times that position in the vector is selected.

```{r}
set.seed(4134)
(choose.these.wr <- srswr(3,6)) # with replacement
(choose.these.wor <- srswor(3,6)) # without replacement
```

Then the `getdata()` function is used to extract the values from the population the indicated number of times.

```{r}
getdata(my.animals, choose.these.wr)
getdata(my.animals, choose.these.wor)
```

This method can be advantageous when drawing with replacement from a large dataset.

## Implementation

::: callout-important
**How to draw a SRS**

1.  Generate a list of all observational units in the population (sampling frame).
2.  Assign each observational unit a unique number, from 1 to the size of the sampling frame $N$.
3.  Use a computer to draw $n$ numbers from 1 to $N$ without replacement.\
4.  Subset the data to keep only the selected rows.
:::

::: {.callout-tip icon="true"}
### Example: Enrollment in Graduate Programs in Statistics

The [last time we saw this example](https://sampling-458.netlify.app/complete_notes/cn03-statistical_foundations_complete.html#variability-of-estimates), we used the `sample()` function to draw 10 values from the single variable/column `FTGradEnrollment`. Let's revisit this data and take an SRS of the entire dataset.

```{r}
stat.phd <- readr::read_csv(here::here("data", "StatisticsPhD.csv")) 
```
:::

**Using the `sample()` function**

[Get in the habit of opening the data set and visually looking at your process at each step. The best way to learn is to check that you know exactly what was done at each step.]{.aside} Generate list of numbers for each observational unit, then draw 10 numbers without replacement.

```{r}
sampling.frame <- 1:NROW(stat.phd) #1, 2
(get.these <- sample(sampling.frame, 10, replace=FALSE)) #3
```

Create a subset data frame with only the rows that were chosen and stored in the vector `get.these`.

```{r}
(sample.stat.phd <- stat.phd[get.these, ])  #4
```

**Using the `srswor()` function**

```{r}
select.these <- srswor(10, NROW(stat.phd))
getdata(stat.phd, select.these)
```

# Estimation under SRS framework

Below is a table of common statistics and how they are estimated under the SRS framework.

| Measure    | Unbiased Estimate $(\hat{\theta})$         | Estimated variance of $(\hat{\theta})$                             |
|------------------|-------------------------|-----------------------------|
| Mean       | $\bar{y} = \frac{1}{n}\sum_{i\in S} y_{i}$ | $\hat{V}(\bar{y}) = (1-\frac{n}{N})\frac{s^{2}}{n}$                |
| Total      | $\hat{\tau} = N\bar{y}$                    | $\hat{V}(\hat{\tau}) = N^{2}\hat{V}(\bar{y})$                      |
| Proportion | $\hat{p} = \bar{y}$                        | $\hat{V}(\hat{p}) = (1-\frac{n}{N})\frac{\hat{p}(1-\hat{p})}{n-1}$ |

[This table can also be found on the [formulas](https://sampling-458.netlify.app/notes/formulas.html) page.]{.aside}

* $i \in S$ : Unit $i$ is an element in the sample $S$

::: callout-important
### Finite Population Correction

$$(1-\frac{n}{N})$$

The larger % of the population that you include in your sample (sampling fraction = $\frac{n}{N}$), the closer you are to a census, the smaller the variability your estimate will have.
:::

* Most samples that are taken from a very large population, the fpc is close to 1.
    * So the variance is more determined by the size of the sample, not the % of the population sampled.

:::{.callout-tip icon=true}
### Example: 
perhaps? we've done a lot of these so far. 
:::

# Sampling Weights 


# Using the `survey` package.

# Determining sample size



# Model based theory vs Randomization based theory (Lohr 2.10, 2.11)

# Closing: When should SRS be used

-   comparison to systemic sampling?
